### 2005-11-14 Report/watershed_paper.tex
Added paper for InsightJournal
darcs-hash:20051114012057-fafb9-6afdc7ac3cd6339c679177e82c0571b49db7ae92.gz

\documentclass{llncs}
\usepackage{graphicx}

\begin{document}
\title{The watershed transform in ITK - discussion and new developments}
\author{Richard Beare{1} \and Gaetan Lehmann{2}}
\institute{Department of Medicine, Moansh University, Australia
\and Biologie du Developpement et de la Reproduction, INRA de Jouy-en-Josas (France)}
\maketitle

\begin{abstract}
This report discusses various definitions and implementations of the
watershed transform as well as providing an introduction to some well
known techniques for applying the watershed transform to practical
problems. The discussion will be focussed on new and existing ITK
classes.
\end{abstract}

\section{Introduction}
The watershed transform is a tool for image segmentation that is fast
and flexible and potentially fairly parameter free. It was originally
derived from a geophysical model of rain falling on a terrain and a variety
of more formal definitions have been devised to allow development of
practical algorithms. A brief discussion of some of these definitions
and algorithms is provided in Section \ref{sect:history}.

Section \ref{sect:usingWT} provides an introduction to a
methodology known as {\em marker based} watershed. Using markers in
watershed transforms is a well established technique in the
morphological image analysis community and has proved very useful for
many applications. The aim of this part of the report is to describe
the technique to a wider audience \footnote{It should also be noted
that the lack of a marker based watershed motivated the recent work on
ITK filter classes.}.

A brief description of two algorithms used to implement the watershed
transform is given in Section \ref{sect:algorithms}.

\section{Image segmentation}
Image segmentation, in the context of this report, may be broadly
defined as finding the part(s) of the image that is of interest in
order to make measurements. This is a basic definition of image
analysis, which is different from image processing and computer
vision\footnote{Image processing produces images, usually for human
consumption, while computer vision aims to produce high level semantic
descriptions of images.}.

It is very rare to achieve a useful segmentation using a single
procedure, even under the most favourable conditions. Successful
segmentation algorithms typically use a carefully constructed
combination of procedures to achieve useful results. Some modern
algorithms, especially the family of routines related to geodesic
active contours, may seem to be an exception to this rule of thumb
because they are often presented as achieving good segmentation
results entirely on their own. However close examination will show
that a variety of preprocessing steps in addition to careful parameter
tuning is necessary to achieve anything useful.

These comments are also relevant to the watershed transform -- it is
very unlikely that any segmentation problem can be solved using the
watershed alone. However it is a very useful and flexible tool that
can be applied in a number of ways. This report will emphasise the
marker based approach to using the watershed transform, but there are
a number of others including watershed trees and region merging
algorithms.

Finally, an important point that should be mentioned, is that the
output of a low level segmentation procedure such as the watershed
transform will typically be a {\em label image}. A label image,
sometimes referred to as a categorical image, has unique values for
each region. For example, if a watershed produces 2 regions, all
pixels belonging to one region would have value $A$, and all belonging
to the other might have value $B$. Unassigned pixels, such as
watershed lines, might have value of $0$.

\section{The watershed transform -- a brief history}
\label{sect:history}
The watershed transform was first proposed by Beucher and
Lantu\'{e}joul\cite{Beucher-Lantu-79} as a geophysical model of rain
falling on a terrain. The idea is that a raindrop falling on a surface
will trickle down the path of steepest descent to a minima. The set of
points on the surface that lead to the same minima is known as a
catchment basin and borders between catchment basins are watershed
lines\footnote{Major geographical features, like rivers or oceans, are
often considered as separated by watershed lines, usually along
mountain ranges.}. If an image is considered as a terrain and divided
into catchment basins then the hope is that each catchment basin would
contain an object of interest. The image representing the terrain will
be referred to as the {\em control surface} in this report.

Translating the model of rain falling on a terrain to an algorithm has
a number of problems\footnote{Although this is what the first ITK
implementation does.} which are discussed in Section
\ref{sect:rainfallmodel}. The alternative is to imagine a flooding
process, with each regional minima\footnote{A regional minima is a set
of pixels surrounded by pixels of higher value.} becoming the source
of a lake. The terrain is progressively flooded and lakes eventually
meet neighboring lakes. Virtual dams are constructed to keep the
neighboring lakes separate as the waterlevel rises. When the image
surface is completely fooded the virtual dams correspond to the
watershed lines\footnote{In 3D neighboring lakes meet at surfaces.}.

A major breakthrough in the implementation of the watershed was made
by Vincent and Soille \cite{Vincent91a} with the introduction of the
first queue based implementation of the watershed transform. This was
especially significant because it allowed similar execution times on
conventional computer systems as was being achieved with specialized
parallel processing systems. The Vincent-Soille algorithm appears to
be used in the matlab image processing toolbox watershed
implementation (based on the observation that the library function has
a {\tt \_vs} suffix).

Another major advance made soon after by Meyer and Beucher
\cite{Beucher93a,Meyer1994a}. Meyer defined the watershed transform in terms of {\em
topographic distances} through {\em lower complete} images. These
definitions rigorously defined the behavior in the presence of
plateaus and lead to some new algorithms based on graph theory. Two
examples of these algorithms are {\em hill-climbing} and {\em image
integration} and will be discussed in more detail in this report.

Readers interested in more technical detail should refer to \cite{Roerdink2000a,Meijster2004a}.

{\em what about the topological watershed?}

\section{Rainfall model}
\label{sect:rainfall}
{\em where should we put this?}  The original concept behind the
watershed transform was rain falling on a terrain and flowing down
paths of steepest descent to local minima. There has been a lot of
work on alternatives to this approach because of the problems with
it. These problems are basically to do with flat zones in an image
which have no paths of steepest descent leading from them, which means
that water would just sit in the one spot. A simple minded
implementation of this idea would therefore produce a separate
catchment basin for each pixel in the flat zone, which is not a very
useful result. The ITK watershed uses a number of steps to avoid this problem:
\begin{enumerate}
\item Label all minima and flat zones.
\item Trace all unlabelled pixels to a labelled one using gradient descent.
\item Check each flat region and find the lowest neighbor. Change the label of the flat region to that of the lowest neighbor.
\end{enumerate}

It should be evident that a number of arbitary choices have been made
here - for example it is impossible to break a flat region into
pieces, which would probably better fit the physical reality of
rainfall (consider a flat topped ridge which should perhaps be split
down the middle). What happens of there are multiple neighboring
regions with equal height - is the labelling decision now somewhat arbitary?

In general the flooding models avoid these dilemas and lead to more
elegant algorithms.

\section{Using the watershed tranform}
\label{sect:usingWT}
\subsection{Classical approach}
The simplest way of using the watershed is to preprocess the image we
want to segment so that the boundaries of our objects are bright (e.g
apply an edge detector) and compute the watershed transform of the
edge image. Watershed lines will correspond to the boundaries and our
problem will be solved. This is rarely useful in practice because
there are always more regional minima than there are objects, either
due to noise or natural variations in the object surfaces. Therefore,
while many watershed lines do lie on significant boundaries, there are
many that don't. Figure \ref{fig:overseg} illustrates the problem. The ITK
watershed implementation provides a way of addressing this problem
using a measure of significance of the boundary and allowing a
threshold to be applied first to reduce the number of local minima.

An Obvious alternative that is very similar in practice is to 
smooth the image so that there are fewer regional minima.
\begin{figure}
\begin{center}
\label{fig:overseg}
\caption{Classical over segmentation.}
\end{center}
\end{figure}

There are some situations where this direct approach can be
useful. The best known of these is probably binary object splitting,
in which the watershed transform is applied to the distance transform
of a mask. This approach is widely used to split touching,
approximately circular objects, such as cells.

\subsection{Watershed from selected minima}
An alternative approach leads us to the concept of markers mentioned
earlier. We will also see that the thresholding approach used by ITK
may be considered as a specific case of the same approach.

Suppose we have preprocessing steps that let select regional minima
that lead to the correct segmentation. How would we modify the
topography of the image so that only the desired regional minima are
present and the watershed lines that remain are the same as they were
before the modification (i.e we only remove the unwanted boundaries)?

This can be achieved by performing a {\em reconstruction by erosion}
or {\em recursive conditional erosion} from the seleted regional
minima. Figure \ref{fig:recon} illustrates the results of apply a
reconstruction by erosion to a 1D image. The original image is shown
in blue and has regional minima A, B, C, D and E. The marker image
which is recursively eroded is shown in black, with a single zero at
regional minimum B. The red plateaus illustrate the parts of the
result image that are above the orginal. The resulting image has only
on regional minima that corresponds to B.
\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=0.5]{Pics/recon.eps}
\label{fig:recon}
\caption{Reconstruction by erosion from a single regional minimum.}
\end{center}
\end{figure}

A more formal definition of recursive erosion is:
\begin{equation}
\label{eq:recon_erosion}
h_{n+1}=\max\{f, \varepsilon h_n\}
\end{equation}
where $h_0(x)$ is the function, with value 0 at the markers and
$\infty$ otherwise and $\varepsilon$ denotes the erosion
operation.

This procedure may not seem to offer much advantage -- after all how
are we going to decide which minima are useful to us? The key step to
using these ideas in practice is to realize that an arbitary mask
image can be used to create minima in an image in locations we want
simply by multiplying the image by the mask. Recursive erosion of the
original image using that mask then produces a topology to which we
can apply a watershed transform. The mask image is commonly called a
marker image.

This is the fundamental concept behind the use of markers in watershed
based segmentation.

Let's summarize the steps again:

\begin{enumerate}
\item Generate a marker image. The marker image will be used to create 
one, possibly extended, minima inside each object of interest. A
marker image will always need to create at least two minima, and at
least one minima will probably be used to segment the background. The
marker image may be generated via interaction with an operator or by
various preprocessing steps.
\item \label{recipe:min}Impose the minima on the control surface. This 
might be done by multiplying by the marker image (or an inverted
version of it, depending on convention).
\item \label{recipe:erode}Recursively erode the control surface to eliminate all minima besides those introduced in the previous step. This produces a new control surface.
\item Apply the watershed tranform to the new surface.
\end{enumerate}

\subsection{Efficiency issues}
The procedure outlined above introduces a lot of additional steps,
some of which could be reasonably computationally expensive
(i.e. recursive erosion). Other steps have also been mentioned in
passing ({\em lower completion} in Section \ref{sect:history} that may
be potentially time consuming. Fortunately we will discover in Section
\ref{sect:algorithms} that the appropriate choice of algorithm implementing the
watershed transform will carry out these steps implicitly.

\section{Algorithms}
\label{sect:algorithms}
The two that will be outlined here are known as {\em image
integration}\cite{Meyer1994a} and {\em hill climbing}
\cite{Beucher93a}, both of which are classed as ordered
algorithms and are examples of graph flooding approaches. These
algorithms implement the watershed tranform by topographic distance
(or approximate it) and are derived from Dijkstra-Moore minimal path
algorithms of graph theory.

\subsection{Ordered queues}
Both of these implementations depend on ordered queues to avoid the
need for explicit computation of lower complete images and recursive
erosion. An ordered queue is a heirarchical priority queue with two
levels. One level is defined by the pixel gray level, with darker
pixels having higher priority, while the second applies to pixels of
the same grey level and provides a FIFO ordering. The FIFO ordering
leads to a breadth first propogation across plateaus because the
ordering becomes related to distance from the regional minima. This is
the mechanism that avoids the to explicitly compute lower slope.

Early implementations used a special ordered queue for 8 bit data,
which may be where the misconception that a morphological watershed
can only be applied to 8 bit data types arose. The new ITK watershed
algorithms use standard c++ containers (maps and standard queues) to
implement the ordered queue\footnote{One potential problem with
applying this style of algorithm to non integer data is finding the
regional minima. Reconstruction approaches are often used to find
regional minima of integer type images, but subtracting 1, carrying
out a recursive dilation and subtracting the two. The same concept
does not extend to floating point data. However, if regional minima
are available the algorithms we are discussing can be applied to
floating point data.}.

In both algorithms each pixel is regarded as a node in a graph that is
connected to neighboring pixels by pairs of edges. Costs of the edges
are defined in terms of a cost related to the gradient.

\subsection{Image integration}
This algorithm maintains a pair of images -- a distance image and a
label image. The distance image is initialized to initialized to
$\infty$ everywhere, except at the minima where it is initialized to
the value of the minima while the label image is initialized to $0$
everywhere, except at the minima where it is initialized to the index
of the minima (or, more typically, the value of the marker image at
those points).

Wavefronts are propogated from each minima, using the ordered queue
approach. As the wavefront reaches pixel the distance travelled by the
wavefront is compared to the distance stored in the distance image. If
the distance travelled by the new wavefront is less than the stored
value then that pixel is given the label associated with the
wavefront, otherwise the label is left unchanged.

\subsection{Hill climbing}
Hill climbing is an approximation that eliminates the need to maintain
a distance image. The approximation employed is that distances between
all neighbors in a 4, 6, 8, 18 or 26\footnote{4 and 8 connected
neighborhoods are used in 2D, 6, 18 and 26 in 3D} and the center of
the neighborhood are equal -- i.e the various $\sqrt{2}$ terms are
ignored. The position of the watershed lines can have some dependence
on raster ordering because of this. In general the approach is very useful and efficient.

Hill climbing requires a label image which is initialized to the
indexes of the regional minima at the regional minima and an UNKNOWN
value elsewhere\footnote{In the ITK implementation a separate image is
used to track UNKNOWN locations because UNKNOWN values are difficult
to deal with in a type independent way.}. The priority queue is
initialised with the borders of the regional minima. Pixels pushed
onto the queue with priority defined by their grey level and labelled
as they are pushed. When pixels are popped from the queue their
neighbors are checked, labelled, and pushed onto the queue if they
aren't watershed pixels. The lower completion step is carried out
implicitly by not allowing pixels to be pushed onto the queue with a
priority lower than the priority of the neighbor which was popped from
the queue.

The assumption that the distance between a pixel and all of its
neighbors is 1 means that the steepest upper neighbors of a pixel are
simply the unlabelled neighbors.

\section{Using watersheds in ITK}

\section{Timing and performance}
streaming and threading.

\section{Discussion}
There is a level of erroneous folklore surrounding the watershed
transform and it is hoped that this report helps to clear the air.

\bibliographystyle{splncs}
\bibliography{Refnames}

\end{document}